---
title: "cps09mar"
author: "jordon griffin"
date: "2023-08-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
```

## Visualizing the earnings distribution

There is no ready-made earnings variable in the CPS. What we are using
is a combination of several variables depending on, among other things,
whether the worker was salaried or paid hourly.
Let’s create a histogram of the constructed earnings variable.
```{r}
cps_mar <- read_xlsx("week_3/cps09mar.xlsx")
earnings_dist <- ggplot(cps_mar, aes(x=earnings)) +
geom_histogram(aes(y = ..density..), color=1,
fill="white", bins=50) +
labs(title="Figure 1. Distribution of earnings, March 2009 CPS")
print(earnings_dist)
```

```{r}
mean(cps_mar$earnings)
median(cps_mar$earnings)
```

```{r}
ggplot(cps_mar, aes(age)) + geom_bar() +
    labs(title="Figure 2. Age distribution, March 2009 CPS")
```
## Relative frequency in the limit

If the formal definition is not your thing, let’s try simulating the idea. Say the event A is
“Heads” in coin flip, so that SN is the number of heads you get from N flips of a fair
coin. We’ll simulate 1000 random trials (flips) of a fair (P(Heads) = p = .5) coin, and
plot the path of the cumulative proportion of heads.
```{r}
# set.seed(12345)
N = 1000
p = .5
flips = sample(x=c(0,1), prob = c(1-p,p), size=N, replace=T)
S_N = cumsum(flips)
N = 1:N
prop_S = S_N/N

flip_data <- data.frame(run=1:1000, prop=prop_S)
flip_plot <- ggplot(flip_data, aes(x=run, y=prop, frame=run)) +
    geom_path(aes(cumulative=T), color="red") +
    xlim(1,1000) + ylim(0.0,1.0) +
    geom_hline(yintercept = 0.5) +
    labs(title="Figure 1. Flipping a fair coin") +
    ylab("Proportion of Heads") +
    xlab("Number of Flips")
print(flip_plot)
```
What is the likelihood of earning $100,000? How can we use the CPS to
infer the answer? One way is to estimate the likelihood by the share of
respondents with incomes greater than or equal to $100,00
```{r}
six_figs <- subset(cps_mar, earnings >= 100000)
six_figs_shr <- nrow(six_figs)/nrow(cps_mar)
print(six_figs_shr)
```

```{r}
earnings_lnorm <- ggplot(cps_mar, aes(earnings)) +
    geom_histogram(aes(y = after_stat(density)), color=1,
        fill="white", bins=50) +
    stat_function(fun = dlnorm,
        args = list(meanlog = mean(log(cps_mar$earnings)),
            sdlog = sd(log(cps_mar$earnings))),
        colour = "red") +
labs(title="Figure 2. Wage histogram with lognormal overlay")
print(earnings_lnorm)
```

```{r}
1-plnorm(100000, meanlog =
    mean(log(cps_mar$earnings)), sdlog = sd(log(cps_mar$earnings)))
```

```{r}
ggplot(cps_mar, aes(x = earnings, y = log(earnings))) +
    geom_point() +
    ggtitle("Relationship between earnings and log earnings") +
    xlab("Earnings") +
    ylab("Log Earnings") +
    geom_segment(aes(x = 0, y = log(100000),
            xend = 100000, yend = log(100000)),
        color = "red", linetype = "dashed") +
    geom_segment(aes(x = 100000, y = 0,
            xend = 100000, yend = log(100000)),
        color = "red", linetype = "dashed") +
    scale_x_continuous(breaks = c(0, 100000), labels = c(0, 100000)) +
    scale_y_continuous(breaks = c(0, log(100000)),
            labels = c(0, round(log(100000), 2)))
```

# Percentage changes
```{r}
y1 <- 105000 # Earnings this year
y0 <- 100000 # Earnings last year
pct_raise_actual <- ((y1-y0)/y0)*100
pct_raise_diflog <- (log(y1)-log(y0))*100
print(paste("Actual percentage raise =", pct_raise_actual))
print(paste("Approx percentage raise =", pct_raise_diflog))
```

# Earning six figures, women vs men

```{r}
six_figs_fvm <- cps_mar %>%
group_by(gender) %>%
summarise(six_figs_shrs = mean(earnings >= 100000))
print(six_figs_fvm)
```