---
title: "BUSN 5000 :: Project I"
subtitle: Exploring Earnings Differences between Women and Men 
author: "Jordon, Griffin"
output:
  html_document:
    self_contained: yes
    toc: true
#    number_sections: true
    theme: readable
#  pdf_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
# Attach packages
library(tidyverse)     # To use the tidy packages
library(modelsummary)  # To make pretty tables
library(here)          # To reference relative file paths
library(ggpmisc)       # To add estimated model annotations

# Set global options
knitr::opts_chunk$set(
  eval=TRUE,
  echo = TRUE, 
  warning = FALSE,
  message = FALSE
  )

# Turn off scientific notation
options(scipen=999)

```

### Preliminaries

#### Academic honesty statement

I have been academically honest in all of my work and will not tolerate academic dishonesty of others, consistent with [UGA's Academic Honesty Policy](https://honesty.uga.edu/Academic-Honesty-Policy/).

Sign the academic honesty statement by typing your name on the **Signature** line. 

**Signature**: Jordon Griffin

We will not accept submissions that omit a signed Academic Honesty statement. 

#### Setting things up 

Before you begin work on this assignment, you should create a new R Project in a selected directory on your computer and adopt the file structure we have promoted in class. We have posted a video on eLC to guide you in this process. 

The `setup` chunk beginning on line 15 lists the packages required for the project. Make sure you have installed them before you begin.

### Instructions and guidance

#### Submission

The `.Rmd` source for this document will be the template for your project submission. Insert your first and last name in `author` component  in line 3 of the `yaml` block. Submit your completed project document as a single PDF. Uploading the PDF to eLC by **1159p on Monday, Oct 2** using the filename `sectiontime_lastname_firstname_projectI.pdf` (e.g. `935_cornwell_chris_projectI.pdf`). 

#### Code chunk evaluation

Each code chunk has some missing elements that you are directed to add. The default setting for `eval= ` is `FALSE` so that you can incrementally turn on the evaluation of code chunks as you complete them. Change `eval= ` to `TRUE` in the code chunk heading to turn evaluation on. When you have completed the entire assignment, all of the chunks should have been evaluated and their results displayed in the knitted output. 

#### Your write-up

Each section instructs you to explain or describe some task or output. Your content should be written exclusively between the green horizontal rules we have inserted into the template. Do *NOT* change the template.

---

### Introduction

<span style="color:green"> 
Provide a brief overview of the project, including: (1) what are you trying to learn about; (2) the data you are using to learn; and (3) a brief summary of your findings. Answer in 3-4 clear and concise sentences. 
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

This project aims to explore the differences in earnings between women and men. We utilize a dataset that contains variables such as age, education, earnings, hours worked, and other relevant factors. Our preliminary exploration indicates several factors that might influence the earnings gap, including education and work hours. Further analyses will delve deeper into these relationships to provide a comprehensive understanding.
<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

---

### Data fundamentals


#### March 2022 CPS

<span style="color:green">
Using the ASEC documentation posted on eLC, provide a brief description of the March 2022 CPS that includes: (1) a  description of the standard monthly CPS; (2) how many households are typically surveyed; (3) what additional information is collected in the ASEC. Answer in 3-4 clear and concise sentences.  
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

The March 2022 Current Population Survey (CPS) is the U.S. Government's official source for employment and unemployment statistics, and it has been conducted monthly for over five decades. Each month, about 54,000 scientifically selected households, representing the nation as a whole, are interviewed. This approach enables monthly and yearly comparisons, minimizing the inconvenience to any particular household. In addition to capturing employment data, the CPS serves as a rich source of demographic details such as age, sex, race, marital status, and educational attainment. Occasionally, supplemental questions on health, education, income, and previous work experience are also included, providing data that helps update the decennial census information.

The Annual Social and Economic (ASEC) Supplement, a part of the CPS, adds further depth. While it contains the basic monthly demographic and labor force data, the ASEC also offers insights on work experience, income types, noncash benefits, health insurance coverage, and migratory patterns. The CPS sample covers the civilian noninstitutional population of the U.S., spanning across 1,328 counties and cities, and includes approximately 100,000 persons aged 15 and above. In addition, the ASEC covers military personnel living in civilian settings and has been supplemented with a sample of Hispanic households since 1976. In total, the ASEC surveys around 95,000 households.<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

```{r read_data, eval =FALSE}
# Read extract created by cpsmar_t.R
cpsmar_t <- read_csv(here("project_1", "cpsmar_t.csv")) 
```

#### March 2022 extract

<span style="color:green">
Explain the actions that you applied to the March 2022 survey in the R script `cpsmar_t.R`, including: (1) a summary of variables created; (2) variables selected; and (3) the number of observations. Answer in 3-4 clear and concise sentences.  
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

In the R script titled cpsmar_t.R, the March 2022 Current Population Survey data undergoes a series of transformations. Two primary datasets, pppub22.csv and hhpub22.csv, are ingested, which consist of individual responses and household data respectively. The household dataset specifically extracts variables related to region and household sequence. A core component of the script calculates the Labor Force Participation Rate (LFPR) for March 2022. This process involves the creation and manipulation of several variables, namely the number of employed individuals, the unemployed, the total labor force, those not in the labor force, the overall population, and the LFPR. These new variables are derived from the major labor force recode "PEMLR" and are conditioned based on age and employment status. After these transformations, the script exports the LFPR values into a text file. The hhpub22.csv dataset comprises 89,197 observations, while the pppub22.csv dataset has 152,732 observations.
<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

--- 

### Beginning to learn

#### From extract to analysis sample

<span style="color:green">
Complete the the `btl1` code chunk by transforming earnings into log earnings and creating a character-valued gender variable.
</span>

```{r btl1, eval=FALSE}
cpsmar_a <- cpsmar_t %>%
  filter(age >= 23, 
         age <= 62,
         earnings>0
         ) %>%
  mutate(learnings = log(earnings),
         gender = ifelse(female == 1, "Female", "Male") # Placeholder for actual transformation logic
         )
```

<span style="color:green">
In 1-2 clear and concise sentences, describe the actions performed in `btl1`.  
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

Do not edit the html formatting. Just delete these instructions and insert your content here. 

<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

#### The baseline earnings distribution

<span style="color:green">
Complete the the `btl2` code chunk by selecting an appropriate `fun` value to model the histogram and providing your Figure 1 with an title.
</span>

```{r btl2, eval=FALSE}
ggplot(cpsmar_a, aes(earnings)) + 
  geom_histogram(
    aes(y = after_stat(density)), color=1, 
    fill="white", bins=50
    ) +
  stat_function(
    fun = dlnorm, 
    args = list(
      meanlog = mean(log(cpsmar_a$earnings)), 
      sdlog   = sd(log(cpsmar_a$earnings))
      ), 
    colour = "red"
    ) +
  labs(
    title="Figure 1. ",
    x="Earnings",
    y="Density"
    )
```

<span style="color:green">
Describe Figure 1, distinguishing the histogram from the red curve overlay and interpreting its general shape. 
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

Do not edit the html formatting. Just delete these instructions and insert your content here. 

<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

#### Summary statistics on earnings, age and gender composition

<span style="color:green">
Complete the `btl3` code chunk by adding `female` to the list of variables and `Median` to the list of items to report, and adding a title to Table 1.
</span>

```{r btl3, eval=FALSE}
datasummary(earnings + age ~ N + Mean,
  data=cpsmar_a,
  title="Table 1. ")
```

<span style="color:green">
Describe the empirical facts reported in Table 1. Answer in 1-2 clear and concise sentences.
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

Do not edit the html formatting. Just delete these instructions and insert your content here. 

<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

#### Earnings distributions for women and men

<span style="color:green">
Complete the `btl4` code chunk, filling in the missing `aes` information and providing a title for Figure 2.
</span>

```{r btl4, eval = FALSE}
ggplot(cpsmar_a, aes(x=_____, group = _____, fill = _____)) + 
  geom_density(adjust=1.5, alpha = 0.4) +
  labs(
    title="Figure 2. ",
    x="Earnings",
    y="Density"
    )
```

<span style="color:green">
In 1-2 clear and concise sentences, describe the main takeaways from Figure 2.
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

Do not edit the html formatting. Just delete these instructions and insert your content here. 

<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

#### Summary statistics on earnings and age by gender

<span style="color:green">
Complete the `btl5` code chunk to create a table of summary statistics (sample size, mean, and median) for earnings and age by gender. Provide a title for Table 2.
</span>

```{r btl5, eval=FALSE}
datasummary( ~ gender*(),
  data=cpsmar_a,
  title="Table 2. ")
```

<span style="color:green">
Describe the empirical facts reported in Table 2 using 1-2 clear and concise sentences.
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

Do not edit the html formatting. Just delete these instructions and insert your content here. 

<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

--- 

### Models for exploration

#### Estimating the CEF, $E(earnings|age)$ 

<span style="color:green">
Complete the `mfe1` code chunk to estimate $E(earnings|age)$ using the analysis sample. "Center" age on 23. Print the 1st, 5th, 10th, ..., 30th estimated conditional mean.
</span>

```{r mfe1, eval=FALSE}
cef <- cpsmar_a %>%
  mutate(age=age- ) %>%
  group_by( ) %>%
  summarise(earnings_bar = mean( ))
print(cef[c(1, 5, 10, , , 30), ])
```

<span style="color:green">
Describe the shape of the estimated CEF and suggest a model for it. Use the printed estimated CEF values to explain how average earnings evolve over a typical career. Limit your response to 2-3 clear and concise sentences.
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

Do not edit the html formatting. Just delete these instructions and insert your content here. 

<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

#### Plotting and modeling the CEF 

<span style="color:green">
Complete the `mfe2` code chunk to plot $\hat{E}(earnings|age)$ with a fit of the model you suggested.
</span>

```{r mfe2, eval=FALSE}
formula <- y~poly(x, , raw=TRUE)
ggplot(data = cef, aes(x= , y= )) + 
  geom_line() + 
  geom_point() +
  geom_smooth(
    formula= , 
    method="lm", 
    se=FALSE, 
    color="red"
    ) +
  stat_poly_eq(
    mapping = use_label(c("eq")), 
    formula= , color = "red"
    ) +
  labs(
    title="Figure 3.  ",
    x="Year",
    y="Average earnings"
    )
```

<span style="color:green">
Use the fitted model to predict earnings in year 0, 14, and 29. Compare these predictions with the estimated conditional means you reported above in 1-2 clear and concise sentences.
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

Do not edit the html formatting. Just delete these instructions and insert your content here. 

<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

#### Gender gap in earnings over a career

<span style="color:green">
Complete the `mfe3` code chunk to estimate $E(earnings|age, gender)$ and plot the results with a fit of an appropriate model. (Note your work on `mfe2` should be helpful here.)
</span>

```{r mfe3, eval=FALSE}
cef_fvm <- cpsmar_a %>%
  mutate( ) %>%
  group_by( ) %>%
  summarise(earnings_bar = mean( ))

formula <- y~poly(x, , raw=TRUE)
ggplot(cef_fvm, aes( ,  ,  )) +
  geom_point() +
  geom_line() +
  geom_smooth(
    
    
    
    
    ) +
  stat_poly_eq(
    aes(label =  paste(stat(eq.label))), 
    formula = formula,
    parse = TRUE
    ) +
  labs(
    title="Figure 5. Log earnings profiles by gender",
    x="Year",
    y="Average log earnings"
    )
```

<span style="color:green">
Describe how the fitted profiles differ and how the gender gap evolves over a typical career. Use the fitted models to compute the male-female ratio in predicted earnings in year 0, 14, 29. Limit your response to 2-3 clear and concise sentences.
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

Do not edit the html formatting. Just delete these instructions and insert your content here. 

<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->

#### Plotting and fitting the percentage gender gap

<span style="color:green">
Complete the `mfe4` code chunk to plot the ratio of male to female earnings to show the how the percentage gap evolves over a typical career. Add a linear fit to the ratio plot.
</span>

```{r mfe4, eval=FALSE}
males    <- filter(cef_fvm, gender == " ") 
females  <- filter(cef_fvm, gender == " ")   
df_ratio <- data.frame(age = males$age, 
                       ratio = males$ /females$ )

ggplot(df_ratio, aes(x =  , y =  )) +
  geom_line() +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  stat_poly_eq(mapping = use_label(c("eq")), color = "blue") +
  labs(
    title="Male-female earnings ratio by age",
    x="Year",
    y="Male-female average earnings ratio"
    )
```

<span style="color:green">
Report the ratio at year 0 and its peak value along with the year it peaks.  Report the predicted gap in each of these years using the linear fit. According to the linear fit, explain by how much the gap is predicted to increase each year. Limit your response to 3-4 clear and concise sentences.
</span>

<!----------------------------------------------------------------------------->
<hr style="border: 1px solid green;">

Do not edit the html formatting. Just delete these instructions and insert your content here. 

<hr style="border: 1px solid green;">
<!----------------------------------------------------------------------------->
